---
title: "ESM 204 Hw 4"
author: "Taylor Medina, Matt McCafferty"
date: '2022-05-16'
output:
  html_document: 
    code_folding: hide
    toc: yes
    theme: paper
---

```{r setup, include = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(here)
library(patchwork)
```

```{r}
damages <- read_csv(here('data/damages.csv'))
warming <- read_csv(here('data/warming.csv'))
```

## Question 1

```{r}
# create quadratic model
damages$warming2 <- damages$warming ^ 2
damages.qm <- lm(damages ~ 0 + warming + warming2, data = damages)

warming_seq <- seq(0, max(damages$warming), by = 0.01)
fit <- predict(damages.qm, list(warming = warming_seq, warming2 = warming_seq ^ 2)) %>% 
  cbind(warming_seq) %>% 
  as.data.frame()

ggplot() +
  geom_point(data = damages, aes(x = warming, y = damages), color = 'grey') +
  geom_line(data = fit, aes(x = warming_seq, y = .), color = 'red') +
  labs(x = expression('warming (°C)'),
       y = 'Total climate damage ($)') +
  theme_minimal()
```

**Figure 1:** Fitted quadratic regression model to environmental damages from warming.

The quadratic regression model modeling environmental damages as a function of global warming is given as:

$$ \hat D = 1.941 \times 10^{13} T ^ 2 - 1.060 \times 10^{12} T $$

## Question 2

```{r}
warming_predict <- warming %>% 
  mutate(damage_np = predict(damages.qm, list(warming = warming_baseline, warming2 = warming_baseline^2)), # damage without pulse
         damage_wp = predict(damages.qm, list(warming = warming_pulse, warming2 = warming_pulse^2)), # damage with pulse
         damages_diff = damage_wp - damage_np, # differences in damage
         new_difference_norm = (damage_wp / 35e9) - (damage_np / 35e9),
         norm_damage_wp = damage_wp / 35e9) # normalized difference

p1 <- ggplot(warming_predict) +
  geom_line(aes(x = year, y = damage_np)) +
  geom_label(aes(2035, 1.8e14, label = 'No Pulse')) +
  labs(y = 'environmental damages ($)') +
  theme_minimal()

p2 <- ggplot(warming_predict) +
  geom_line(aes(x = year, y = damage_wp)) +
  geom_label(aes(2035, 1.8e14, label = 'Pulse')) +
  labs(y = 'environmental damages ($)') +
  theme_minimal()

p3 <- ggplot(warming_predict) +
  geom_line(aes(x = year, y = damages_diff)) +
  labs(y = 'environmental damage \n difference ($)') +
  theme_minimal()

p4 <- ggplot(warming_predict) +
  geom_line(aes(x = year, y = new_difference_norm)) +
  labs(y = 'environmenta damage \n difference ($/ton)') +
  theme_minimal()

(p1 | p2) / (p3 | p4)

```

**Figure 2:** Plots of damages from no pulse (top-left), with the 35b ton pulse (top-right), the differences in damages (with pulse - no pulse) (bottom-left), and the normalized differences in damages per ton of carbon (bottom-right.)

## Question 3

The present discounted value formula is given as the following:

$$ PV = \sum_{i = 0}^{78}{\frac{D_i}{(1+r)^i}}$$
This takes into account that 2022 is considered year 0 (i = 0) and 2100 is the final value (i = 78). $D_i$ is the damage associated with the year.

```{r}
# function to calculate present value based off of one r
pv <- function(data, r){
  
  pv.df <- data.frame()
  
  for (i in 1:length(data)){ # r starts index at 1
  
     pv <- data[i] / (1 + r) ^ (i - 1) # making i = 0 for the first iteration
     pv.df <- rbind(pv.df, pv)
     col_sum <- colSums(pv.df)
     
  }
  
  return(col_sum)
  
}

# function creates a dataframe for multiple r values
by_r <- function(data, r){
  
  df <- data.frame()
  
  for(i in 1:length(r)){
    
    pv_r <- pv(data, r[i])
    df <- rbind(df, pv_r)
    
  }
  
  final.df <- cbind(r, df)
  return(final.df)
  
}

rseq <- seq(0.02, 0.03, by = 0.001) 
r.df <- by_r(warming_predict$norm_damage_wp, rseq) 
colnames(r.df) <- c("discount_rate", "SCC")

ggplot(r.df) +
  geom_line(aes(x = discount_rate, y = SCC)) +
  geom_point(aes(x = 0.021, y = r.df$SCC[2]), color = 'purple') +
  geom_segment(aes(x = 0.021, y = 30000, xend = 0.021, yend = r.df$SCC[2]), linetype = 'dashed', color = 'purple') +
  geom_label(aes(0.0217, 40000, label = 'r = 0.021')) +
  labs(y = "Social Cost of Carbon ($/ton)", x = "Discount Rate") +
  theme_minimal()
```

**Figure 3:** Social Cost of Carbon as a function of the discount rate. The optimal discount from the Ramsey Rule is shown with the purple dashed line. 

## Question 4

The Ramsey Rule is defined as the following:

$$ r = \rho + \eta g $$
The optimal discount rate given $\rho$ = 0.001, $\eta$ = 2, and g = 0.01 is:

$$ r = 0.001 + 2(0.01) $$

$$ \boxed{r = 0.021}$$

## Question 5

Policy A: Warming will occur at baseline (p = 0.5) or 1.5 * baseline (p = 0.5)

Policy B: warming will occur at baseline until 2050 and stay that way forever (1.29°C).

r = 0.02 and society is risk neutral

```{r}
policy <- warming %>% 
  mutate(a_1 = predict(damages.qm, list(warming = warming_baseline, warming2 = warming_baseline ^ 2)), # Policy A baseline
         a_2 = predict(damages.qm, list(warming = 1.5 * warming_baseline, warming2 = (1.5 * warming_baseline) ^ 2)), # 1.5 * policy A
         expected_a = 0.5 * a_1 + 0.5 * a_2, # E[A]
         b_warming = ifelse(warming_baseline > 1.29, 1.29, warming_baseline), # modifies warming so it levels off after 1.29 C
         expected_b = predict(damages.qm, list(warming = b_warming, warming2 = b_warming ^ 2))) # E[B]

pv_a <- pv(policy$expected_a, 0.02) %>% 
  rbind(pv_b <- pv(policy$expected_b, 0.02)) 
```

The expected present value of Policy A is $`r pv_a[1]`.

The expected present value of Policy B is $`r pv_a[2]`.

```{r}
x <- seq(0, 5e15, by = 1e12)

costs <- data.frame(x = x) %>% 
  mutate(a = pv_a[1],
         b = pv_a[2] + x)

b.lm <- lm(b ~ x, data = costs)

eq <- (costs$a[1] - b.lm$coefficients[1]) / b.lm$coefficients[2]

ggplot(costs) +
  geom_line(aes(x = x, y = a), color = 'red') +
  geom_line(aes(x = x, y = b)) +
  geom_segment(aes(x = eq, y = 0, xend = eq, yend = 6e15), linetype = 'dashed', color = 'purple') +
  geom_label(aes(1e15, 6e15, label = 'Policy B')) +
  geom_label(aes(3.5e15, 6e15, label = 'Policy A')) +
  labs(x = 'Additional Costs ($)',
       y = 'Total Present Value ($)') +
  theme_minimal()
```

**Figure 4:** Plot of Total Present Value vs. Additional Costs for Policy A (red line) and Policy B (black line). The purple vertical dashed line is the boundary where Policy B costs more than policy A.

When additional costs for Policy B < `r eq`, the Policy B is cheaper. When the additional costs > `r eq`, then policy A is cheaper.

## Question 6
```{r}

```

If society is risk-averse, then they will prefer the expected value with the most certainty. This would be Policy B since there is less uncertainty with that policy than policy A.


